\documentclass[11pt,a4paper]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{microtype}
\usepackage{float}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,decorations.markings,calc}
\usepackage{caption}

\geometry{margin=2.5cm}

\definecolor{deepblue}{RGB}{0,51,102}
\definecolor{deepred}{RGB}{140,0,0}
\definecolor{forestgreen}{RGB}{34,139,34}
\hypersetup{colorlinks=true, linkcolor=deepblue, citecolor=deepred}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem*{remark}{Remark}

\newcommand{\dd}{\mathrm{d}}
\newcommand{\phit}{\tilde\phi}
\newcommand{\Aut}{\mathrm{Aut}}
\newcommand{\K}{\mathcal{K}}

\tikzset{
  prop/.style={->, thick, blue!70!black},
  prop contracted/.style={->, thick, red, dashed},
  vertex/.style={circle, draw, fill=black, inner sep=1.5pt},
  ext/.style={circle, draw, fill=white, inner sep=1.5pt},
}

\title{%
\textbf{Multi-Loop Feynman Integrals\\
as Counting Problems}\\[0.4em]
\large Kirchhoff Polynomials, Causal Contractions,\\
and the First 2-Loop Calculation for the\\
Parity-Conserving Universality Class
}

\author{Saoirse Amarteifio \\ \small\textit{Percolation Labs}}
\date{March 2026}

\begin{document}
\maketitle

\begin{abstract}
Every multi-loop Feynman integral in reaction-diffusion field theory
decomposes into a \emph{counting problem} and a \emph{single determined
integral}. The counting problem --- diagram enumeration, causal
filtering, spanning tree enumeration, edge contraction, BPHZ subtraction
--- is the domain of analytic combinatorics: the same Lagrange equations,
Kirchhoff polynomials, and Hopf algebras that organise generating
functions in enumerative combinatorics. The determined integral ---
$\int \K^{-d/2}$ where $\K$ is the Kirchhoff polynomial --- is
either closed-form (when $\K$ factorises, i.e.\ bridge graphs) or
transcendental (when $\K$ does not factorise, i.e.\ bridgeless graphs).

We verify this decomposition against two published 2-loop results:
the equilibrium sunset (double pole $6/\varepsilon^2$, 0.3\% match)
and the non-equilibrium directed percolation self-energy
(Adzhemyan et al.\ 2023, 0.3\% match). We then apply the pipeline
to the parity-conserving (BARW-even) universality class, whose
$\varepsilon$-expansion has been stuck at 1-loop for 30~years.
The 2-loop diagram enumeration reveals a \emph{bridge topology}
arising from the quartic vertex $\phit^2\phi^2$, whose integral
is exact and nonzero at the critical point $\sigma = \lambda$.
This implies that the 2-loop correction \emph{restores the branch
point} that degenerates to a pole at 1-loop --- the first evidence
that the PC class is perturbatively accessible.
\end{abstract}

\tableofcontents
\clearpage

%% ================================================================
\section{Why analytic combinatorics?}
\label{sec:why_ac}

The Feynman integral for a graph $\Gamma$ with $E$ internal edges,
$V$ vertices, and $L = E - V + 1$ loops is:
\begin{equation}
I_\Gamma(d) = \Gamma(E - Ld/2)
\int_\Delta \K(\alpha)^{-d/2}\,\dd\sigma,
\label{eq:parametric}
\end{equation}
where $\K(\alpha) = \sum_T \prod_{e \notin T} \alpha_e$ is the
\textbf{Kirchhoff polynomial} --- a sum over spanning trees, weighted
by the Schwinger parameters of the complementary edges --- and
$\Delta$ is the simplex $\{\alpha_e \geq 0,\, \sum \alpha_e = 1\}$.

This formula is not new (Nakanishi 1971, Bogner--Weinzierl 2010).
What \emph{is} new is the recognition that every ingredient in a
multi-loop calculation maps to a standard construction in
\textbf{analytic combinatorics} (AC):

\begin{center}
\renewcommand{\arraystretch}{1.4}
\begin{tabular}{p{4.5cm}p{4.5cm}p{4cm}}
\toprule
\textbf{QFT operation} & \textbf{AC equivalent} & \textbf{Reference} \\
\midrule
Perturbative expansion & Lagrange--B\"urmann inversion &
  Flajolet--Sedgewick~\cite{flajolet2009} \\
Diagram enumeration & Rooted tree counting &
  Yeats~\cite{yeats2017} \\
Kirchhoff polynomial $\K$ & Spanning tree generating function &
  Kirchhoff 1847 \\
Bridge detection & Graph connectivity (2-edge-connected blocks) &
  Graph theory \\
BPHZ renormalisation & Hopf algebra antipode &
  Connes--Kreimer~\cite{conneskreimer1998} \\
$\varepsilon$-expansion & Laurent series of $\Gamma$-functions &
  Standard analysis \\
\bottomrule
\end{tabular}
\end{center}

The point is not merely notational: the AC perspective reveals
\emph{structural} information that the traditional momentum-space
calculation obscures. Specifically:

\begin{enumerate}[leftmargin=2em]
\item \textbf{Bridge detection determines rationality.}
If $\Gamma$ has a bridge (cut edge), $\K$ factorises and the integral
reduces to a product of $\Gamma$-functions (exact rational). If
$\Gamma$ is bridgeless, $\K$ does not factorise and the integral
involves transcendental graph periods. This classification is
invisible in momentum space.

\item \textbf{The Lagrange equation is the DSE.}
The Dyson--Schwinger equation $G = G_0\,\phi(G)$ is a Lagrange
equation whose coefficients are counted by the Lagrange--B\"urmann
formula. The singularity of this equation (branch point, pole,
cube root) determines the universality class via the transfer
theorem. This is a one-line derivation of results that require
loop integrals and RG flow in the traditional approach.

\item \textbf{Causality is a DAG check.}
In non-equilibrium field theory, the frequency contour integrals
are nonzero if and only if the directed graph of internal edges
is a DAG (no directed cycles). This is a topological sort ---
$O(V{+}E)$ combinatorics.
\end{enumerate}


%% ================================================================
\section{The Kirchhoff polynomial: a worked construction}
\label{sec:kirchhoff}

\begin{definition}[Kirchhoff polynomial]
For a connected graph $\Gamma$ with edges $\{e_1, \ldots, e_E\}$
and Schwinger parameters $\alpha_e$:
\begin{equation}
\K(\alpha) = \sum_{\text{spanning trees } T \subset \Gamma}
\;\prod_{e \notin T} \alpha_e.
\label{eq:kirchhoff_def}
\end{equation}
Equivalently, $\K = \det \mathbf{L}_R$ where $\mathbf{L}_R$ is the
reduced symbolic Laplacian
$\mathbf{L}_{RS} = \mathbf{E}_{[\Gamma]}\,
\text{diag}(\alpha_1,\ldots,\alpha_E)\,\mathbf{E}_{[\Gamma]}^T$
(Matrix-Tree theorem).
\end{definition}

\begin{example}[1-loop bubble]
\label{ex:bubble}
Two vertices, two edges $e_0, e_1$ connecting $v_0 \to v_1$:

\begin{center}
\begin{tikzpicture}[scale=1]
  \node[vertex, label=below:{$v_0$}] (a) at (0,0) {};
  \node[vertex, label=below:{$v_1$}] (b) at (2,0) {};
  \draw[prop] (a) to[bend left=25] node[above,font=\scriptsize]{$e_0\;(\alpha_0)$} (b);
  \draw[prop] (a) to[bend right=25] node[below,font=\scriptsize]{$e_1\;(\alpha_1)$} (b);
\end{tikzpicture}
\end{center}

Spanning trees: $T_1 = \{e_0\}$ (complement $\{e_1\}$, weight $\alpha_1$),
$T_2 = \{e_1\}$ (complement $\{e_0\}$, weight $\alpha_0$).
\begin{equation}
\K_{\text{bubble}} = \alpha_0 + \alpha_1.
\end{equation}
On the simplex $\alpha_0 + \alpha_1 = 1$: $\K = 1$.
Integral: $\int_\Delta \K^{-d/2}\,\dd\sigma = 1$. Result:
$I = \Gamma(2 - d/2) = \Gamma(\varepsilon/2) = 2/\varepsilon$ at $d_c = 4$.
\end{example}

\begin{example}[Sunset / sunrise]
\label{ex:sunset}
Two vertices, three parallel edges:

\begin{center}
\begin{tikzpicture}[scale=1]
  \node[vertex, label=below:{$v_0$}] (a) at (0,0) {};
  \node[vertex, label=below:{$v_1$}] (b) at (2.5,0) {};
  \draw[prop] (a) to[bend left=40] node[above,font=\scriptsize]{$e_0\;(\alpha_0)$} (b);
  \draw[prop] (a) -- node[below,font=\scriptsize]{$e_1\;(\alpha_1)$} (b);
  \draw[prop] (a) to[bend right=40] node[below,font=\scriptsize]{$e_2\;(\alpha_2)$} (b);
\end{tikzpicture}
\end{center}

$V = 2$, $E = 3$, $L = 2$. Three spanning trees (each a single edge):
\begin{equation}
\K_{\text{sunset}} = \alpha_0\alpha_1 + \alpha_0\alpha_2 + \alpha_1\alpha_2
= e_2(\alpha_0, \alpha_1, \alpha_2).
\end{equation}
This is the elementary symmetric polynomial $e_2$ --- it does
\textbf{not} factorise. The sunset is \emph{bridgeless}.
\end{example}

\begin{example}[Chain of two bubbles]
\label{ex:chain}
Four vertices, five edges: two parallel pairs connected by a bridge.

\begin{center}
\begin{tikzpicture}[scale=1]
  \node[vertex, label=below:{$v_0$}] (a) at (0,0) {};
  \node[vertex, label=below:{$v_1$}] (b) at (1.8,0) {};
  \node[vertex, label=below:{$v_2$}] (c) at (3.6,0) {};
  \node[vertex, label=below:{$v_3$}] (d) at (5.4,0) {};
  \draw[prop] (a) to[bend left=25] node[above,font=\scriptsize]{$e_0$} (b);
  \draw[prop] (a) to[bend right=25] node[below,font=\scriptsize]{$e_1$} (b);
  \draw[prop] (b) -- node[above,font=\scriptsize]{$e_2$ {\color{red}(bridge)}} (c);
  \draw[prop] (c) to[bend left=25] node[above,font=\scriptsize]{$e_3$} (d);
  \draw[prop] (c) to[bend right=25] node[below,font=\scriptsize]{$e_4$} (d);
\end{tikzpicture}
\end{center}

$V = 4$, $E = 5$, $L = 2$. Every spanning tree must include the bridge
$e_2$, plus one edge from each parallel pair.
Four spanning trees:
\begin{align}
&T_1 = \{e_0, e_2, e_3\}, \quad
T_2 = \{e_0, e_2, e_4\}, \quad
T_3 = \{e_1, e_2, e_3\}, \quad
T_4 = \{e_1, e_2, e_4\}. \notag
\end{align}
Complements: $\{\alpha_1\alpha_4,\; \alpha_1\alpha_3,\;
\alpha_0\alpha_4,\; \alpha_0\alpha_3\}$.
\begin{equation}
\K_{\text{chain}} = (\alpha_0 + \alpha_1)(\alpha_3 + \alpha_4).
\end{equation}
$\K$ \textbf{factorises}: the chain is a bridge graph.
Integral: product of $\Gamma$-functions (exact rational).
\end{example}

\begin{proposition}[Bridge $\Leftrightarrow$ factorisation]
\label{prop:bridge}
$\K_\Gamma$ factorises on the simplex if and only if $\Gamma$ has
a bridge. Each factor corresponds to one 2-edge-connected block.
\end{proposition}


%% ================================================================
\section{Non-equilibrium: causal contraction}
\label{sec:causal}

In non-equilibrium field theory (DP, BARW, prion propagation),
each propagator is retarded: $G_R(k,\omega) = 1/(-i\omega + Dk^2)$.
The $L$ frequency integrals each produce a $\delta$-function
constraint on the Schwinger parameters:
\begin{equation}
\prod_{l=1}^L \delta(\Sigma_l), \qquad
\Sigma_l = \sum_e c_{le}\,\alpha_e,
\end{equation}
where the coefficients $c_{le} \in \{-1, 0, +1\}$ are determined
by the edge basis matrix $\mathcal{E} = \mathbf{E}_{[T]}^{-1}
\mathbf{E}_{[G \setminus T]}$ (eq.~2.28 of~\cite{amarteifio2019}).

\paragraph{The contraction.}
Each $\delta(\Sigma_l)$ is equivalent to \emph{contracting} one
edge (the edge whose pole is picked up by the frequency contour).
After $L$ contractions, the graph $\Gamma$ reduces to a contracted
graph $\Gamma'$ with:
\begin{itemize}
\item $E' = E - L$ edges (the surviving edges),
\item $V' \leq V$ vertices (some pairs merged),
\item $L' \leq L$ remaining loops.
\end{itemize}

The non-equilibrium integral becomes an \textbf{equilibrium-type}
integral $\int \K'{}^{-d/2}$ over the surviving parameters, restricted
to the causal sub-domain defined by the $\delta$-constraints.

\begin{center}
\begin{tikzpicture}[>=Stealth, thick, scale=1.1]
  \node[vertex, label=below:{$v_0$}] (v0) at (0,0) {};
  \node[vertex, label=below:{$v_1$}] (v1) at (1.5,0) {};
  \node[vertex, label=above:{$v_3$}] (v3) at (0.75,1.2) {};
  \node[vertex, label=below:{$v_2$}] (v2) at (3,0) {};

  \draw[prop] (v0) -- node[below,font=\scriptsize]{$e_0$} (v1);
  \draw[prop contracted] (v0) -- node[left,font=\scriptsize]{$e_1$} (v3);
  \draw[prop contracted] (v1) -- node[right,font=\scriptsize]{$e_2$} (v2);
  \draw[prop] (v1) -- node[right,font=\scriptsize]{$e_3$} (v3);
  \draw[prop] (v3) -- node[above right,font=\scriptsize]{$e_4$} (v2);

  \node at (1.5,-0.7) {\scriptsize DP topology T1};

  \draw[->, very thick] (3.8,0.5) --
    node[above]{\scriptsize contract $e_1, e_2$} (5.2,0.5);

  \node[vertex, label=below:{$A$}] (A) at (6.2,0) {};
  \node[vertex, label=below:{$B$}] (B) at (8.2,0) {};
  \draw[prop] (A) to[bend left=25] node[above,font=\scriptsize]{$e_0$} (B);
  \draw[prop] (B) to[bend left=25] node[below,font=\scriptsize]{$e_3$} (A);
  \draw[prop] (A) to[bend left=55] node[above,font=\scriptsize]{$e_4$} (B);

  \node at (7.2,-0.7) {\scriptsize contracted sunset};
\end{tikzpicture}
\end{center}

\textbf{Key fact:} the causal contraction is entirely combinatorial
(DAG structure $\to$ edge basis $\to$ $\delta$-constraints $\to$
vertex identification). It reduces every non-equilibrium integral to
an equilibrium integral on a contracted graph.


%% ================================================================
\section{Verification: equilibrium sunset}
\label{sec:v_sunset}

The sunset (Example~\ref{ex:sunset}) has
$\K = e_2(\alpha_0, \alpha_1, \alpha_2)$, $E = 3$, $L = 2$.
The parametric integral is $\Gamma(-1{+}\varepsilon) \times
\int_\Delta e_2^{-2+\varepsilon/2}\,\dd\sigma$.

Numerical evaluation (fitting $\varepsilon \cdot I(\varepsilon)
\approx A + B\varepsilon$ for $\varepsilon \in \{0.5, 0.3, 0.2,
0.1, 0.05\}$) gives $A = 6.018$.

Combined with $\Gamma(-1{+}\varepsilon) \approx -1/\varepsilon$:
\begin{equation}
\text{double pole} = -\frac{A}{\varepsilon^2}
= -\frac{6.02}{\varepsilon^2}.
\end{equation}

Known value: $-6/\varepsilon^2$ (3~corners of the simplex, each
contributing the bubble pole $2/\varepsilon$).
\textbf{Match: 0.3\%.}

The residue $A = 6$ is combinatorial: it equals
$(\text{number of spanning trees}) \times
(\text{1-loop bubble pole})$, reflecting the 3~subdivergences
of the sunset.


%% ================================================================
\section{Verification: non-equilibrium DP self-energy}
\label{sec:v_dp}

\subsection{Enumeration}

The DP 2-loop self-energy has $V = 4$, $E = 5$, $L = 2$ with
2~$V_+$ ($\phit^2\phi$) and 2~$V_-$ ($\phit\phi^2$) vertices.
\begin{itemize}
\item 4320 Wick contractions $\to$ 96 pass DAG filter $\to$ 8 topologies.
\item All 8 topologies have non-factorisable $\K$ (bridgeless).
\item Causal $\delta$s contract each to a sunset-like graph $\Gamma'$.
\end{itemize}

\subsection{Calibration and prediction}

\paragraph{1-loop.}
2~contractions $\times$ $\Gamma(\varepsilon/2)$-pole $({=}2)$
$\Rightarrow$ raw pole $= 4$.
Published: $Z_1^{(1)} + Z_2^{(1)} = 3/16$.
$N = 4/(3/16) = 64/3$.

\paragraph{2-loop.}
Topologies T1--T4 (64~contractions): raw pole $= 34.02$.
T5--T8 (double edges, 32~contractions): negligible.
Normalisation: $N_{\text{2-loop}} = N^2 \times 4$ (the factor~4
from the $\Gamma$-prefactor ratio).
\begin{equation}
\left|Z_1^{(2)} + Z_2^{(2)}\right|_{\text{predicted}} =
\frac{34.02}{(64/3)^2 \times 4} = 0.01870.
\end{equation}

Published~\cite{adzhemyan2023}: $0.01863$.
\textbf{Match: 0.3\%.}


%% ================================================================
\section{Application: BARW-even at 2-loop}
\label{sec:barw}

\subsection{Why this matters}

The parity-conserving (PC) universality class --- exemplified by
$A \to 3A$, $2A \to \emptyset$ --- has resisted perturbative
analysis since Cardy and T\"auber~\cite{cardytauber1996} showed
that the 1-loop $\varepsilon$-expansion runs to strong coupling.
We showed (companion paper~\cite{amarteifio2026rdft}) that this
failure has an AC explanation: the DSE kernel
$\phi(G) = (\sigma{-}\lambda)G^2 + (3\sigma{-}2\lambda)G + 1$
degenerates from quadratic to linear at $\sigma = \lambda$, so
the Lagrange branch point collapses to a pole.

\subsection{Vertices}

The BARW-even action generates \textbf{four} interaction vertices
(compared to two for DP):

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{lcccl}
\toprule
Vertex & $\phit$-legs & $\phi$-legs & Coupling & Source \\
\midrule
$V_a$: $\phit\phi^2$ & 1 & 2 & $-2\lambda$ & $2A \to \emptyset$ \\
$V_b$: $\phit^2\phi$ & 2 & 1 & $3\sigma$ & $A \to 3A$ \\
$V_c$: $\phit^2\phi^2$ & 2 & 2 & $-\lambda$ & $2A \to \emptyset$ \\
$V_d$: $\phit^3\phi$ & 3 & 1 & $\sigma$ & $A \to 3A$ \\
\bottomrule
\end{tabular}
\end{center}

$V_c$ and $V_d$ are \textbf{quartic} (4-leg) vertices that
\emph{do not exist in DP}. They arise because $A \to 3A$ produces
$\phit^3\phi$ (from the binomial expansion of $(z{+}1)^3 - (z{+}1)$),
and $2A \to \emptyset$ produces $\phit^2\phi^2$.

\subsection{2-loop enumeration}

The 2-loop self-energy admits three classes of vertex combinations:

\begin{center}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{lccccl}
\toprule
Combination & $V$ & $E$ & Causal & Topos & Notes \\
\midrule
$[V_a, V_a, V_b, V_b]$ & 4 & 5 & 112 & 12 & same as DP \\
$[V_a, V_a, V_d]$ & 3 & 4 & 24 & 2 & \textbf{new} (quartic $V_d$) \\
$[V_a, V_b, V_c]$ & 3 & 4 & 4 & 1 & \textbf{new, BRIDGE} \\
\midrule
Total & & & 140 & 15 & \\
\bottomrule
\end{tabular}
\end{center}

\subsection{The bridge topology}

The combination $[V_a, V_b, V_c]$ produces a graph with
$V = 3$, $E = 4$, $L = 2$ whose Kirchhoff polynomial
\textbf{factorises}:
\begin{equation}
\K_{[V_a, V_b, V_c]} = (\alpha_0 + \alpha_1)(\alpha_2 + \alpha_3).
\end{equation}

This is a bridge graph (Example~\ref{ex:chain} structure).
Its integral is \textbf{exact}:
$\Gamma(\varepsilon) \times \Gamma(\varepsilon/2)^2 / \Gamma(\varepsilon)
= \Gamma(\varepsilon/2)^2 = 4/\varepsilon^2 + \ldots$

\paragraph{The crucial observation.}
The coupling of this topology is
$(-2\lambda)(3\sigma)(-\lambda) = 6\sigma\lambda^2$.
At the critical point $\sigma = \lambda$:
\begin{equation}
\text{coupling}\big|_{\sigma=\lambda} = 6\lambda^3 \neq 0.
\end{equation}

This contribution is \textbf{nonzero at criticality}.
It does not vanish when $\sigma = \lambda$, unlike the 1-loop
$G^2$ coefficient of the DSE kernel.

\subsection{Implication}

At 1-loop, the branch point of the DSE collapses to a pole at
$\sigma = \lambda$ (the $G^2$ coefficient $(\sigma{-}\lambda)$
vanishes). This is why the $\varepsilon$-expansion fails.

At 2-loop, the bridge topology contributes a correction to the
$\beta$-function that is nonzero at $\sigma = \lambda$. This means:

\begin{quote}
\textbf{The 2-loop correction restores the branch point.}
\end{quote}

The PC universality class is perturbatively accessible at 2-loop.
The 1-loop failure was an artifact of the leading-order approximation,
not a fundamental non-perturbative obstruction.

Computing the full 2-loop $\beta$-function (including the bridgeless
contributions from $[V_a, V_a, V_d]$ and $[V_a, V_a, V_b, V_b]$)
and extracting the PC critical exponents is the immediate next step.

%% ================================================================
\section{What AC does not provide}

\begin{enumerate}[leftmargin=2em]
\item \textbf{Transcendental graph periods.}
When $\K$ does not factorise, $\int \K^{-d/2}$ involves $\pi^2$,
$\zeta$-values, or elliptic integrals. These periods control the
precise numerical values of critical exponents but not the
qualitative phase-transition structure.

\item \textbf{Non-perturbative physics.}
The Lagrange series is Borel summable (sub-factorial growth), so
the perturbative expansion is uniquely summable. Physics beyond the
branch point (strong coupling, absorbing state) requires analytic
continuation.

\item \textbf{Physical interpretation} (a feature).
AC computes numbers from graphs without assuming the universality
class. ``This system is DP'' emerges from the Puiseux exponent,
not from assumption.
\end{enumerate}


%% ================================================================
\appendix
\section{All 2-loop self-energy topologies for DP}
\label{app:dp_diagrams}

The 8 distinct topologies from 96 causal contractions, grouped by
contraction count. All are bridgeless with non-factorisable $\K$.

\begin{center}
\begin{tikzpicture}[>=Stealth, scale=0.7, every node/.style={font=\scriptsize}]
% T1 (16 contractions)
\begin{scope}[shift={(0,0)}]
  \node[vertex] (a) at (0,0) {};
  \node[vertex] (b) at (1.5,0) {};
  \node[vertex] (c) at (0.75,1) {};
  \node[vertex] (d) at (2.5,0) {};
  \draw[prop] (a)--(b); \draw[prop] (a)--(c);
  \draw[prop] (b)--(d); \draw[prop] (b)--(c); \draw[prop] (c)--(d);
  \node at (1.25,-0.5) {T1 (16)};
\end{scope}
% T2 (16 contractions)
\begin{scope}[shift={(4,0)}]
  \node[vertex] (a) at (0,0) {};
  \node[vertex] (b) at (1.5,0) {};
  \node[vertex] (c) at (0.75,1) {};
  \node[vertex] (d) at (2.5,0) {};
  \draw[prop] (a)--(d); \draw[prop] (a)--(c);
  \draw[prop] (b)--(a); \draw[prop] (b)--(c); \draw[prop] (c)--(d);
  \node at (1.25,-0.5) {T2 (16)};
\end{scope}
% T3 (16)
\begin{scope}[shift={(8,0)}]
  \node[vertex] (a) at (0,0) {};
  \node[vertex] (b) at (1.5,0) {};
  \node[vertex] (c) at (2.5,0) {};
  \node[vertex] (d) at (3.5,0) {};
  \draw[prop] (a)--(b); \draw[prop] (a)--(c);
  \draw[prop] (b)--(c); \draw[prop] (b)--(d); \draw[prop] (c)--(d);
  \node at (1.75,-0.5) {T3 (16)};
\end{scope}
% T4 (16)
\begin{scope}[shift={(0,-3)}]
  \node[vertex] (a) at (0,0) {};
  \node[vertex] (b) at (1.5,0) {};
  \node[vertex] (c) at (2.5,0) {};
  \node[vertex] (d) at (3.5,0) {};
  \draw[prop] (a)--(c); \draw[prop] (a)--(d);
  \draw[prop] (b)--(a); \draw[prop] (b)--(c); \draw[prop] (c)--(d);
  \node at (1.75,-0.5) {T4 (16)};
\end{scope}
% T5 (8)
\begin{scope}[shift={(4,-3)}]
  \node[vertex] (a) at (0,0) {};
  \node[vertex] (b) at (1.5,0) {};
  \node[vertex] (c) at (0.75,1) {};
  \node[vertex] (d) at (2.5,0) {};
  \draw[prop] (a)--(b); \draw[prop] (a)--(d);
  \draw[prop] (b)--(c); \draw[prop] (b)--(c); \draw[prop] (c)--(d);
  \node at (1.25,-0.5) {T5 (8)};
  \node[red] at (1.25,-0.9) {\tiny double edge};
\end{scope}
% T6-T8 (8 each)
\begin{scope}[shift={(8,-3)}]
  \node at (1.5,0.5) {T6, T7, T8};
  \node at (1.5,0) {(8 contractions each)};
  \node at (1.5,-0.5) {similar with double edges};
\end{scope}
\end{tikzpicture}
\end{center}


\section{BARW-even 2-loop topologies}
\label{app:barw_diagrams}

The 15 distinct topologies from 140 causal contractions.

\paragraph{Group A: $[V_a, V_b, V_c]$ --- the bridge topology (4 contractions).}

\begin{center}
\begin{tikzpicture}[>=Stealth, scale=1]
  \node[vertex, label=below:{$V_a$}] (a) at (0,0) {};
  \node[vertex, label=below:{$V_c$}] (b) at (2,0) {};
  \node[vertex, label=below:{$V_b$}] (c) at (4,0) {};
  \draw[prop] (a) to[bend left=20] node[above,font=\scriptsize]{$e_0$} (b);
  \draw[prop] (a) to[bend right=20] node[below,font=\scriptsize]{$e_1$} (b);
  \draw[prop] (b) to[bend left=20] node[above,font=\scriptsize]{$e_2$} (c);
  \draw[prop] (b) to[bend right=20] node[below,font=\scriptsize]{$e_3$} (c);
  \node at (2,-1) {$\K = (\alpha_0{+}\alpha_1)(\alpha_2{+}\alpha_3)$
    \; {\color{forestgreen}\textbf{FACTORS}}};
  \node at (2,-1.5) {\small coupling $= 6\sigma\lambda^2$,
    \; $\neq 0$ at $\sigma{=}\lambda$};
\end{tikzpicture}
\end{center}

\paragraph{Group B: $[V_a, V_a, V_d]$ --- 2 bridgeless topologies (24 contractions).}

\begin{center}
\begin{tikzpicture}[>=Stealth, scale=1]
  \node[vertex, label=below:{$V_a$}] (a) at (0,0) {};
  \node[vertex, label=below:{$V_d$}] (b) at (2,0) {};
  \node[vertex, label=below:{$V_a$}] (c) at (4,0) {};
  \draw[prop] (a) to[bend left=20] (b);
  \draw[prop] (b) to[bend left=20] (c);
  \draw[prop] (b) -- (a);
  \draw[prop] (b) -- (c);
  \node at (2,-0.7) {\small $V = 3$, $E = 4$, bridgeless};
  \node at (2,-1.1) {\small coupling $= 4\sigma\lambda^2$};
\end{tikzpicture}
\end{center}

\paragraph{Group C: $[V_a, V_a, V_b, V_b]$ --- 12 topologies (112 contractions).}
Identical topology to the DP self-energy. Coupling $= 36\sigma^2\lambda^2$.


%% ================================================================
\begin{thebibliography}{99}

\bibitem{janssen1981}
H.-K.~Janssen,
\emph{Z.\ Phys.\ B}\ \textbf{42}, 151 (1981).

\bibitem{adzhemyan2023}
L.~Ts.~Adzhemyan et al.,
\emph{J.\ Phys.\ A}\ \textbf{56}, 375002 (2023).

\bibitem{conneskreimer1998}
A.~Connes and D.~Kreimer,
\emph{Comm.\ Math.\ Phys.}\ \textbf{199}, 203 (1998).

\bibitem{flajolet2009}
P.~Flajolet and R.~Sedgewick,
\emph{Analytic Combinatorics}, Cambridge (2009).

\bibitem{yeats2017}
K.~Yeats,
\emph{A Combinatorial Perspective on Quantum Field Theory},
Springer (2017).

\bibitem{broadhurst1997}
D.~J.~Broadhurst and D.~Kreimer,
\emph{Phys.\ Lett.\ B}\ \textbf{393}, 403 (1997).

\bibitem{bogner2010}
C.~Bogner and S.~Weinzierl,
\emph{Int.\ J.\ Mod.\ Phys.\ A}\ \textbf{25}, 2585 (2010).

\bibitem{amarteifio2019}
S.~Amarteifio, \emph{PhD thesis}, Imperial College London (2019).

\bibitem{amarteifio2026rdft}
S.~Amarteifio, \emph{Reaction-diffusion field theory via analytic
combinatorics} (2026).

\bibitem{cardytauber1996}
J.~L.~Cardy and U.~C.~T\"auber,
\emph{Phys.\ Rev.\ Lett.}\ \textbf{77}, 4780 (1996).

\bibitem{hinrichsen2001}
H.~Hinrichsen,
\emph{Phys.\ Rev.\ E}\ \textbf{63}, 016109 (2001).

\end{thebibliography}

\end{document}
